Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2022-12-21 BulletinOfMathBio/sn-article.tex
add BMB script

\documentclass{article}
% \documentclass{sn-jnl}
% \documentclass[review]{elsarticle}

\usepackage{arxiv}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}   % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage[square,compress,numbers]{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{float}
\usepackage{lineno}
% \linenumbers

% \def\bibcommenthead{}%

\newcommand{\insertPdfFig}[3]{
  \begin{figure}[H]
  \centering
  \includegraphics[width=8.4cm]{#1.pdf}
  \caption{#2}
  \label{fig:#1}
  \end{figure}
}

\newcommand{\insertSmallPdfFig}[3]{
  \begin{figure}[h]
  \centering
  \includegraphics[width=8.4cm]{#1.pdf}
  \caption{#2}
  \label{fig:#1}
  \end{figure}
}

% \title{On the nature of the Gompertz in relation to the epidemic in 2019-2021}
\title{Unifying the communicable disease spreading paradigm with Gompertzian growth}
% \graphicspath{{../figures/}}
%\date{September 9, 1985} % Here you can change the date presented in the paper title
%\date{}          % Or removing it

\author{Matz A. Haugen, \\
Independent Reseacher, Oslo\\
\texttt{matzhaugen@gmail.com} \\
  \And
  Dorothea Gilbert \\
  Independent Reseacher, Oslo\\
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{On the nature of the Gompertz}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={A template for the arxiv style},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={},
pdfkeywords={},
}

\begin{document}
\maketitle

\begin{abstract}
A number of studies have shown that cumulative mortality followed a Gompertz curve in the initial Covid pandemic period, March-April 2020. 
We show that the Gompertz curve is incompatible with expected initial logistic growth curves as predicted by traditional Susceptible-Infected-Recovered (SIR) models, and propose a new theory which better explains the nature of the mortality characteristics based on a global biosphere disturbance. 
Second, we show that for the Gompertz curve to emerge, the disturbance has to act on everyone simultaneously, rejecting the possibility of a disease propagation stage. 
Third, we connect logistic growth with Gompertzian growth by augmenting the logistic growth equation with higher order interaction terms, and show that the SIR model family is compatible with Gompertzian growth only when all nodes in the transmission network communicate with infinite speed and interaction. Crucially, this augmentation must be accompanied by a causality-reversal where the source of growth is not the pool of infected but the pool of susceptible people. We thus find a novel bridge between logistic and Gompertzian growth, separate from the existing Richards model (also called $\theta$-logistic growth).
\end{abstract}

% keywords can be removed
\keywords{Gompertz \and Coherence \and Covid \and Coronavirus \and Network Analysis \and Stochastic}

\section*{Introduction}
Traditional communicable disease spreading theory assumes a pathogen which infects the population through a network of transmission. 
Following this line of reasoning it can be theoretically shown that in the early stages of an epidemic, growth follows a logistic-like curve, where the very beginning exhibits exponential growth, and for which analytic and semi-analytic solutions have recently emerged \cite{harko2014exact,kroger2020analytical,schlickeiser2021analytical,heng2020approximately}. 
A large body of research reveals how these models fit the mortality seen due to the Covid pandemic in 2020, a pandemic due to an influenza-like disease \cite{carletti2020covid,cooper2020sir,postnikov2020estimation,munoz2021sir,cooper2022dynamical,saikia2021covid}.

@@

\section*{Unifying the SIR model with Gompertzian growth}
It is possible to reconcile the SIR model with the Gompertz curve under a common model different from the traditional Richards model. First, start with the networked SIR equation for the infected group,
\begin{equation}
\label{eq:SIRnetwork}
\frac{d x_i}{dt} = \beta (1-x_i)\sum_j{a_{ij}}x_j - \alpha x_i\quad \forall i.
\end{equation}

@@

\subsection*{Code Availability}

Scripts used to produce the plots are downloadable in the form of a Jupyter notebook using Python from 
\newline
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/gompertz_vs_sird.ipynb}

\subsection*{Data Availability}
Observations are taken from the Github repository compiled by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, Baltimore,
USA \cite{dong2020interactive}, and are available at \url{https://github.com/CSSEGISandData/COVID-19}, specifically the table \newline\verb|csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv|, \newline downloaded Aug. $5^{th}$ 2022.

@@

@@

\section*{Supplement}

% Estimating the parameters using real data requires some consideration around the variance of the growth variable as a function of time. This variance will in turn inform the estimation process by considering a weighted minimization of the residuals of the observations regressed onto the model. In the stochastic Gompertz process postulated in Section \ref{seq:alt}, the process under a log-transformation is the Ornstein-Uhlenbeck process which has a closed from expression for its transition probability, and thus its variance as a function of time. This variance can then be used in a weighted least squares estimation technique or any other weighted technique. In the original domain there is also a closed form expression available but it is somewhat more convoluted and given that observations in this domain are visually quite homoscedastic with equal variance across time (Fig \ref{fig:Gompertz_vs_SIRD} and Tables \ref{tab:gomp} and \ref{tab:sird}), this is expression and consideration is omitted. But for the log-domain, observations exibit changing variance with time.

% To be explicit, if we consider the notation from Section \ref{seq:alt} and let $P(z, t|z', t')$ be the transition probability from some growth and time coordinate $z', t'$ to some other growth and time coordinate $z, t$, than the transition probability of the Olstein-Uhlenbeck process in \ref{eq:microscopic} is given by\cite{risken1996fokker},
% \begin{equation}
% P(z, t|z', t') = \sqrt{\frac{\beta}{2\pi \sigma (1 - e^{-2\beta (t-t')})}}\exp{\left[ -\frac{\beta (z - e^{-\beta(t-t')}z')^2}{2\sigma(1-e^{-2 \beta(t-t')})} \right]}.
% \end{equation}

% If we consider $t$ the time at which the stationary distribution is reached and that this stationary distribution is given as a delta function, variance of the growth will be given as
% \begin{equation}
% \text{Var[Z(t)|t']} = \sigma(1-e^{-2 \beta(t-t')}).
% \end{equation}

% Because weights on observations in a regression are invariant to multiplication we can let $\sigma=1$. The slope coefficient would have to be estimated either by a standard least squares or through the 3-parameter model in (\ref{eq:GompertzODE}).

\insertPdfFig{Gompertz_vs_SIRD}{Cumulative diseased plotted against number of days after $Y(t) / Y_{max}>0.005$, comparing an SIRD with a Gompertz model. Both models are fit using non-linear least squares according to equations (\ref{eq:GompertzODE}) and (\ref{eq:SIRD_D}) in the text. Fitting is done with Python-Scipy and a 6-day windowed average of deaths. Observations are taken from the Github repository compiled by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, Baltimore, USA \cite{dong2020interactive}.}

% \insertPdfFig{Gompertz_vs_SIRD_gompfit_log_domain}{Log transformed cumulative diseased plotted against its time derivative after $Y(t) / Y_{max}>0.005$, with a straight line fit regressing onto the $50^{th}$ quantile.}

%  Carletti (left) transformation to our notation (right)
% gamma - eta
% eta - xi
% beta - kappa 
\begin{table}
\label{tab:sird}
\centering
\caption{Estimated values for the SIRD model fitted to various countries' observations as shown in Figure 1 using notation from eq. (12). Standard deviation estimates is given in the parenthesis corresponding to the last significant digit of the reported number.}
\begin{tabular}{llll}
\toprule
{} &           $\eta\times 10^{-1}$ &      $\xi \times 10^{4}$ &            $\kappa\times 10^{2}$ \\
\midrule
Belgium        &    94(8) &   3.2(2) &   9.3(7) \\
Portugal       &   6.6(4) &    35(2) &   4.5(3) \\
Austria        &   6.9(5) &    40(3) &   8.7(6) \\
Switzerland    &    22(1) &  13.6(7) &  11.0(5) \\
Sweden         &    19(5) &     8(2) &     3(1) \\
Italy          &   137(3) &  1.77(5) &  3.71(1) \\
France         &  240(10) &  1.24(7) &   8.1(4) \\
Spain          &   145(2) &  2.27(5) &  4.75(8) \\
Germany        &    61(4) &   3.9(3) &   6.5(5) \\
Denmark        &   2.7(2) &    96(8) &   4.3(3) \\
United Kingdom &   181(5) &  1.39(5) &   4.3(1) \\
Netherlands    &    29(1) &   8.9(5) &   4.3(2) \\
Turkey         &    34(3) &   6.8(5) &   6.9(6) \\
Iran           &  22.3(4) &  10.4(4) &  2.51(7) \\
US             &   312(5) &  0.76(2) &  1.83(7) \\
China          &  130(30) &   3.5(5) &    26(4) \\
\bottomrule
\end{tabular}

\end{table}

\begin{table}
\label{tab:gomp}
\centering
\caption{Estimated values for the Gompertz model fitted to various countries' observations, as shown in Figure 1 using notation from eq. (1).}
\begin{tabular}{llll}
\toprule
{} &             $\nu\times 10$ &           $\beta\times 10^{2}$ &        $\tilde{Y}\times 10^{2}$ \\
\midrule
Belgium        &   9.3(1) &   8.1(2) &  10.1(9) \\
Portugal       &  6.15(3) &  6.22(7) &   7.0(3) \\
Austria        &  6.98(5) &   7.6(1) &   7.8(4) \\
Switzerland    &  8.08(6) &   8.1(1) &   8.9(5) \\
Sweden         &   4.9(1) &   4.3(2) &     6(1) \\
Italy          &  7.60(6) &  5.88(7) &   8.4(4) \\
France         &  10.2(1) &   8.2(1) &  11.0(8) \\
Spain          &  9.41(9) &   7.5(1) &  10.2(7) \\
Germany        &  7.56(6) &  6.53(8) &   8.4(4) \\
Denmark        &  5.94(7) &   6.5(2) &   6.8(7) \\
United Kingdom &  8.46(5) &  6.54(7) &   9.3(4) \\
Netherlands    &  7.34(6) &  6.52(8) &   8.2(4) \\
Turkey         &  7.04(5) &  6.38(8) &   7.9(4) \\
Iran           &  5.87(6) &  5.04(8) &   6.7(4) \\
US             &  7.53(7) &  5.32(7) &   8.4(5) \\
China          &   9.8(1) &   9.4(2) &    10(1) \\
\bottomrule
\end{tabular}

\end{table}

\bibliographystyle{abbrvnat}
\bibliography{references2}  %%% Uncomment this line and comment out the ``thebibliography'' section below to use the external .bib file (using bibtex) .

%%% Uncomment this section and comment out the \bibliography{references} line above to use inline references.
% \begin{thebibliography}{1}

%   \bibitem{kour2014real}
%   George Kour and Raid Saabne.
%   \newblock Real-time segmentation of on-line handwritten arabic script.
%   \newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
%       International Conference on}, pages 417--422. IEEE, 2014.

%   \bibitem{kour2014fast}
%   George Kour and Raid Saabne.
%   \newblock Fast classification of handwritten on-line arabic characters.
%   \newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
%       International Conference of}, pages 312--318. IEEE, 2014.

%   \bibitem{hadash2018estimate}
%   Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
%   Jacovi.
%   \newblock Estimate and replace: A novel approach to integrating deep neural
%   networks with existing applications.
%   \newblock {\em arXiv preprint arXiv:1804.09028}, 2018.

% \end{thebibliography}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
%%\documentclass[default]{sn-jnl}% Default
%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\newcommand{\insertSmallPdfFig}[3]{
  \begin{figure}[h]
  \centering
  \includegraphics[width=8.4cm]{#1.pdf}
  \caption{#2}
  \label{fig:#1}
  \end{figure}
}
\newcommand{\insertPdfFig}[3]{
  \begin{figure}[h]
  \centering
  \includegraphics[width=12.4cm]{#1.pdf}
  \caption{#2}
  \label{fig:#1}
  \end{figure}
}
% \graphicspath{{../figures/}}

\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

\title[Unifying the communicable disease spreading paradigm]{Unifying the communicable disease spreading paradigm with Gompertzian growth}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1]{\fnm{Matz} \sur{Haugen}}\email{matzhaugen@gmail.com}

\author[2]{\fnm{Dorothea} \sur{Gilbert}}\email{D.Thea@protonmail.ch}
% \equalcont{These authors contributed equally to this work.}

\affil*[1]{\orgname{Abel X}, \orgaddress{\city{Oslo}, \postcode{1360}, \state{Baerum}, \country{Norway}}}

\affil[2]{\orgname{Independent Researcher}, \orgaddress{\city{Oslo}, \country{Norway}}}

% \affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}

%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

\abstract{A number of studies have shown that cumulative mortality followed a Gompertz curve in the initial Covid pandemic period, March-April 2020. 
We show that the Gompertz curve is incompatible with expected initial logistic growth curves as predicted by traditional Susceptible-Infected-Recovered (SIR) models, and propose a new theory which better explains the nature of the mortality characteristics based on a global biosphere disturbance. 
Second, we show that for the Gompertz curve to emerge, the disturbance has to act on everyone simultaneously, rejecting the possibility of a disease propagation stage. 
Third, we connect logistic growth with Gompertzian growth by augmenting the logistic growth equation with higher order interaction terms, and show that the SIR model family is compatible with Gompertzian growth only when all nodes in the transmission network communicate with infinite speed and interaction. Crucially, this augmentation must be accompanied by a causality-reversal where the source of growth is not the pool of infected but the pool of susceptible people. We thus find a novel bridge between logistic and Gompertzian growth, separate from the existing Richards model (also called $\theta$-logistic growth).
}

%%================================%%
%% Sample for structured abstract %%
%%================================%%

% \abstract{\textbf{Purpose:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Methods:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Results:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
% 
% \textbf{Conclusion:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}

\keywords{Gompertz, Coherence, Covid, Communicable Diseases, Network, Stochastic Processes}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

\section{Introduction}\label{sec1}
Traditional communicable disease spreading theory assumes a pathogen which infects the population through a network of transmission. 
Following this line of reasoning it can be theoretically shown that in the early stages of an epidemic, growth follows a logistic-like curve, where the very beginning exhibits exponential growth, and for which analytic and semi-analytic solutions have recently emerged \cite{harko2014exact,kroger2020analytical,schlickeiser2021analytical,heng2020approximately}. 
A large body of research reveals how these models fit the mortality seen due to the Covid pandemic in 2020, a pandemic due to an influenza-like disease \cite{carletti2020covid,cooper2020sir,postnikov2020estimation,munoz2021sir,cooper2022dynamical,saikia2021covid}.

@@

\subsection*{Unifying the SIR model with Gompertzian growth}
It is possible to reconcile the SIR model with the Gompertz curve under a common model different from the traditional Richards model. First, start with the networked SIR equation for the infected group,
\begin{equation}
\label{eq:SIRnetwork}
\frac{d x_i}{dt} = \beta (1-x_i)\sum_j{a_{ij}}x_j - \alpha x_i\quad \forall i.
\end{equation}

@@

\backmatter

\bmhead{Supplementary information}

A supplement is available for this article.
% \bmhead{Acknowledgments}

% \section*{Declarations}

% Some journals require declarations to be submitted in a standardised format. Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section. If yes, your manuscript must contain the following sections under the heading `Declarations':

\section*{Competing interests}
The authors have no relevant financial or non-financial interests to disclose.”

\section*{Data Availability}
Observations are taken from the Github repository compiled by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, Baltimore,
USA \cite{dong2020interactive}, and are available at \url{https://github.com/CSSEGISandData/COVID-19}, specifically the table \newline\verb|time_series_covid19_deaths_global.csv|, \newline downloaded Aug. $5^{th}$ 2022.

@@

\section*{Code Availability}
Scripts used to produce the plots are downloadable in the form of a Jupyter notebook using Python from 
\newline
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/gompertz_vs_sird.ipynb}
and \newline
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/hypergeometric.ipynb}

\section*{Funding}
The authors declare that no funds, grants, or other support were received during the preparation of this manuscript.

@@

%%===================================================%%
%% For presentation purpose, we have included        %%
%% \bigskip command. please ignore this.             %%
%%===================================================%%
% \bigskip
% \begin{flushleft}%
% Editorial Policies for:

% \bigskip\noindent
% Springer journals and proceedings: \url{https://www.springer.com/gp/editorial-policies}

% \bigskip\noindent
% Nature Portfolio journals: \url{https://www.nature.com/nature-research/editorial-policies}

% \bigskip\noindent
% \textit{Scientific Reports}: \url{https://www.nature.com/srep/journal-policies/editorial-policies}

% \bigskip\noindent
% BMC journals: \url{https://www.biomedcentral.com/getpublished/editorial-policies}
% \end{flushleft}

% \begin{appendices}

% \section{Section title of first appendix}\label{secA1}

% An appendix contains supplementary information that is not an essential part of the text itself but which may be helpful in providing a more comprehensive understanding of the research problem or it is information that is too cumbersome to be included in the body of the paper.

% %%=============================================%%
% %% For submissions to Nature Portfolio Journals %%
% %% please use the heading ``Extended Data''.   %%
% %%=============================================%%

% %%=============================================================%%
% %% Sample for another appendix section			       %%
% %%=============================================================%%

% %% \section{Example of another appendix section}\label{secA2}%
% %% Appendices may be used for helpful, supporting or essential material that would otherwise 
% %% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
% %% tables and equations etc.

% \end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%

\bibliography{references2}
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

%% Default %%
%%\input sn-sample-bib.tex%

### 2022-12-21 arxiv/main.tex
add BMB script

Scripts used to produce the plots are downloadable in the form of a Jupyter notebook using Python from 
\newline
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/gompertz_vs_sird.ipynb}

@@

\insertPdfFig{Gompertz_vs_SIRD}{Cumulative diseased plotted against number of days after $Y(t) / Y_{max}>0.005$, comparing an SIRD with a Gompertz model. Both models are fit using non-linear least squares according to equations (\ref{eq:GompertzODE}) and (\ref{eq:SIRD_D}) in the text. Fitting is done with Python-Scipy and a 6-day windowed average of deaths. Observations are taken from the Github repository compiled by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, Baltimore, USA \cite{dong2020interactive}.}

>>>>

Scripts used to produce the plots are downloadable in the form of a Jupyter notebook using Python from 
\newline
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/gompertz_vs_sird.ipynb}
and 
\url{https://nbviewer.org/urls/emf-research.fra1.digitaloceanspaces.com/gompertz/hypergeometric.ipynb}

@@

\insertPdfFig{Gompertz_vs_SIRD}{Cumulative diseased plotted against number of days after $Y(t) / Y_{max}>0.005$, comparing an SIRD with a Gompertz model. Both models are fit using non-linear least squares according to equations (\ref{eq:GompertzODE}) and (\ref{eq:SIRD_D}) in the text. Fitting is done with Python-Scipy and a 6-day windowed average of deaths. The basic reproduction number, $R_0$ is computed as specified in the Methods. Observations are taken from the Github repository compiled by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, Baltimore, USA \cite{dong2020interactive}.}
